Estimating demographic bias on tests of children's early vocabulary

### 2022-01-21 paper/cogsci.Rmd
demographic betas x pruning threshold

abstract: >
    Children's early language skill has been linked to later educational outcomes, making
    it important to accurately measure early language. Parent-reported instruments such as the 
    Communicative Development Inventories (CDIs) have been shown to provide valid, consistent 
    measures of children's aggregate early language skill. However, CDIs are predominantly 
    comprised of hundreds of vocabulary items, some of which may not be heard (and thus learned) 
    equally often by children of varying backgrounds. Here, we use a database of American English 
    CDIs to identify words that show strong bias for particular groups of children, on dimensions
    of sex (male vs. female), race (white vs. non-white), and socioeconomic status 
    (high vs. low). For each dimension, we identify dozens of strongly biased items, and show that 
    eliminating these items reduces the expected ability difference between groups. For sex, we
    consider how to propose replacement words that may show less bias, on the basis of their 
    relatively equal frequency in adult speech directed to male and female children.

@@

Researchers, clinicians, and parents have long been fascinated with the surprising speed and variability in the growth of young children's vocabulary.
Children's early vocabulary growth is assumed to reflect not only their exposure to child-directed speech, but also the varying difficulty of different types of words, and individual differences in the aptitude of the child -- including potential language deficits.
Research has uncovered both within-child consistency in development, as well as significant influence from external factors.
For example, @bornstein2016stability found stability in core language skills across 10 years of children's development, despite changes in maternal income and education over the study period.
Yet socioeconomic status (SES) has also been found to be associated with children's language processing and vocabulary by 18 months [@fernald2013ses], and to be predictive of later educational outcomes [@marchman2008speed; see @schwab2016 for a review].
Demographic factors are also predictive of language skill: first-born children tend to outpace their siblings, and female children tend to have better language skills than their age-matched male counterparts [@fenson1994] -- a sex-based verbal advantage that continues through high school [see @petersen2018gender for a review].
However, it is also difficult to get a complete measure of young children's language skill: long recordings are prohibitively difficult to collect and transcribe [see @roy2015predicting, the exception that proves the rule], and yet any short recording (e.g., a 1-hour play session) will elicit only a small proportion of the words and constructions that children know. 
Thus, researchers of early word learning have constructed tests with hundreds of words, with an intentional bias to select words that are more likely to be known by children.

@@

While it is quite reasonable to expect that some CDI words are easier than others, and even to use these varying difficulties to predict variation in children's language ability, the use of psychometric models highlights the possibility that some CDI items may function differently (i.e., be more/less difficult) for different groups of children.
The idea that some items on a test may show bias, favoring one group over another, is known as Differential Item Function (DIF; @holland1993differential]. 
On any given test, it is clearly undesirable to have more items favoring one group (say, children from rural households) over another (urban children), as the test will overestimate the ability of test-takers in the former (rural) group -- and not because of any underlying mean difference in ability between the groups, but simply because the test is unfair [@camilli2013ongoing].
A variety of statistical methods for detecting DIF have been proposed, and investigations have in several instances identified DIF for many items on tests favoring one group over another (e.g., rural vs. urban).
Our goal here is to test the items on the American CDI:WS for DIF along three main axes: sex (male vs. female), socioeconomic status (low- vs. high-SES), and race (white vs. non-white).

<!--
This problem can be be understood through a concrete example that is well-known from research on early language learning: consider the fact that females show a larger vocabulary than males across early development [see @frank2021 Ch. 6?; OTHER REFS]. 
The question is whether or not females actually have a higher ability level, or if the tests predominantly have words that are easier for females (i.e., biased).
(It could even be that the test is biased toward males but the language ability of females is large enough to overcome that bias.)
To address this question, we need to know the ability levels of girls and boys in order to confirm that if a word is learned earlier by girls it is simply because of an ability difference rather than a bias inherent to the word. The problem arises from the fact that we measure ability level using the very same test that we are trying to check for biased words. 
If the boys and girls were of the same average language ability, this would not be an issue but given the evidence showing that girls have a higher ability level, we need to know the difference in ability so that when we find DIF for a specific word can be sure it is outside of the expecting DIF that results naturally from their difference in ability.
-->

@@

```{r, echo=F}
sex_tab <- table(d_demo$sex)
ses_tab <- table(d_demo$ses_group)
eth_tab <- table(d_demo$ethnicity)
eth2_tab <- table(d_demo$eth_group)
```

We analyze parent-reported Wordbank data from `r nrow(d_demo)` American English CDI: Words & Sentences administrations for children `r min(d_demo$age)` to `r max(d_demo$age)` months of age [@frank2017; @frank2021].
Full demographic data are not reported in some datasets contributed to Wordbank: sex was available for `r nrow(d_demo)-sum(is.na(d_demo$sex))` children, race/ethnicity was available for `r nrow(d_demo)-sum(is.na(d_demo$ethnicity))` children, and mother's education (a proxy for socioeconomic status; SES) was available for `r nrow(d_demo)-sum(is.na(d_demo$ses_group))` children.

The analysis of sex-based differences included CDI administrations from `r sex_tab["Female"]` female and `r sex_tab["Male"]` male children.
The analysis of race-based differences included data from `r eth_tab["White"]` white, `r eth_tab["Asian"]` Asian, `r eth_tab["Black"]` Black, `r eth_tab["Hispanic"]` Hispanic, and `r eth_tab["Other"]` "Other" children. 
For this analysis we categorized participants' race/ethnicity as White (`r eth2_tab["White"]`) or Non-white (`r eth2_tab["Nonwhite"]`). 
For the SES-based analysis, we categorized the `r ses_tab["high"]` children whose mothers had at least some college or more as high-SES, and those whose mothers had at most high school (`r ses_tab["low"]` children) as low-SES.

@@

In the multigroup Rasch model, an item's difficulty is allowed to vary depending on which group an individual child belongs to. 
For example, in the sex-based multigroup model, item $i$'s difficulty is $b_{i}^{female}$ for females, and $b_{i}^{male}$ for males.
In the quest to identify DIF, a multigroup Rasch model will be fitted for each demographic dimension of interest (sex, socioeconomic status, and ethnicity).
To contextualize these results, we first fitted a baseline Rasch model to the entire dataset, without demographic information.

@@

@@

Figure 1 shows children's language ability vs. age by demographic group from the baseline Rasch model, which assumes no DIF (i.e., equal item parameters for all groups).
A linear regression for each demographic group, with age (centered) and its interaction, showed significant effects in the expected directions, as can also be seen in the figure. 
Female children had higher language ability than male children ($\beta=0.56$, $p<.001$), with no significant interaction with age ($\beta=0.02$, $p=.10$).
High-SES children had higher language ability than low-SES children ($\beta=0.23$, $p=.02$), an advantage that grew with age ($\beta=0.11$, $p<.001$).
White children had higher language ability than non-white children ($\beta=0.50$, $p<.001$), an advantage that grew with age ($\beta=0.08$, $p<.001$).
We will re-examine these demographic regressions after trimming items showing extreme DIF.

@@

The first step we take to identify CDI items with DIF is to create GLIMMERs (Graphs of Logits Imputed Multiply with Means Equal; @stenhaug2021treading), which visualize between-item variation in group performance differences.
These parameters are drawn from a fitted multigroup Rasch model for each demographic variable (sex, SES, and race), with the assumption that the mean language ability in each group is the same (e.g. for sex, $\mu_{male}=\mu_{female}=0$), thus pushing all between-group variation into the item difficulty parameters.
For the case of sex, where we believe $\mu_{female}>\mu_{male}$, this means we may expect to find many items with difficulty $b_j^{female} < b_j^{male}$, but we can still examine the distribution of differences in item difficulty ($d_j = b_j^{male} - b_j^{female}$) for outliers.
To give analysts a sense of the uncertainty about the existence of DIF, GLIMMER plots show distributions of parameter differences rather than point estimates. 
These distributions are generated by drawing 10,000 imputations from the item parameter covariance matrix.

Figures 2-4 show GLIMMERs for a selection of CDI items for sex, SES, and race, respectively.
The full GLIMMERs, with all 680 CDI:WS items, are available on [OSF](https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a), but were too large to include here.
Nonetheless, it is important to inspect the full plots, for if there is a cluster of items in a GLIMMER, the analyst may conclude that these items are strong candidates for DIF on that dimension.
For example, there is some clustering at the top and bottom of Figure 2: at the top, "vagina", "tights", "dress (object)", and "doll" form a cluster of items that are much easier for females, while at the bottom, "penis" stands out as much easier for males.
For the SES and race GLIMMERs (Figures 3 and 4)--and in the rest of the full sex GLIMMER, there are not clusters, but rather a continuum of smoothly varying differences with overlap. 
This makes identifying items with DIF quite difficult, as different methods are likely to yield inconsistent results [@stenhaug2021treading].
Nonetheless, we describe the distributions of item-level group difficulty differences, and examine overlap of the extrema.

@@

mm_sex <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06
mm_ses <- get_extreme_item_difficulty_differences(mm_ses) # min: -1.07 max: 1.13
mm_eth <- get_extreme_item_difficulty_differences(mm_eth) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex, extreme),
                     subset(mm_ses, extreme),
                     subset(mm_eth, extreme)) 
# length(unique(all_extreme$definition)) 
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

@@

For sex, the median difficulty difference (male-female) was `r round(median(mm_sex$d_diff),2)` (M=`r round(mean(mm_sex$d_diff),2)`, sd=`r round(sd(mm_sex$d_diff),2)`), with `r length(which(mm_sex$d_diff>0))`/680 items being easier for females than males.
The fact that the bulk of this distribution favors females seems to confirm that the female language advantage is pervasive across the CDI:WS, and is thus likely to be a real advantage.
For SES, the median difficulty difference (low-high) was `r round(median(mm_ses$d_diff),2)` (M=`r round(mean(mm_ses$d_diff),2)`, sd=`r round(sd(mm_ses$d_diff),2)`), with `r length(which(mm_ses$d_diff>0))`/680 items being easier for high-SES than low-SES children. 
With the mean and median difficulty differences close to 0, and roughly half of the words favoring each SES group, it is tempting to conclude that the CDI:WS items are somewhat balanced with respect to SES.
For race, the median difficulty difference (nonwhite-white) was `r round(median(mm_eth$d_diff),2)` (M=`r round(mean(mm_eth$d_diff),2)`, sd=`r round(sd(mm_eth$d_diff),2)`), with `r length(which(mm_eth$d_diff>0))`/680 items being easier for white than non-white children, revealing a fairly pervasive advantage for white children on CDI items.

Finally, for each distribution we looked at extreme values, flagging items with difficulty differences that were outside 2 standard deviations.
This identified 25 items with extreme sex-based difficulty differences, only 7 of which were easier for females, with the other 18 items favoring males. <!-- thus, removing these extrema would actually lower male children's vocabulary scores..] -->
For SES, 39 extrema were identified, only 13 of which were easier for low-SES children.
For race, 31 extrema were identified, only 11 of which were easier for non-white children.
Thus, while if anything the bulk of the extrema in the sex-based model advantaged the lower language ability group (males), the extrema in the other two models mostly favored the advantaged groups: 26/39 extreme items in the SES model favored high-SES children, and 20/31 items in the race model favored white children.
17 CDI items were identified as extrema in more than one model: "tractor" and "vroom" were extreme in all three models, "choo choo" was extreme in both SES and sex models, "give me five!" was extreme in both race and sex models, and 13 other items were extreme in both SES and race models ("grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy", "pet's name", "up", and "so").
In our final analysis, we explore the effect of pruning the 15 items that were extreme in both SES and race models.

@@

>>>>

abstract: >
    Children's early language skill has been linked to later educational outcomes, making
    it important to accurately measure early language. Parent-reported instruments such as the 
    Communicative Development Inventories (CDIs) have been shown to provide valid, consistent 
    measures of children's aggregate early language skill. However, CDIs are predominantly 
    comprised of hundreds of vocabulary items, some of which may not be heard (and thus learned) 
    equally often by children of varying backgrounds. Here, we use a database of American English 
    CDIs to identify words that show strong bias for particular groups of children, on dimensions
    of sex (male vs. female), race (white vs. non-white), and maternal education 
    (high vs. low). For each dimension, we identify dozens of strongly biased items, and show that 
    eliminating these items reduces the expected ability difference between groups. For sex, we
    consider how to propose replacement words that may show less bias, on the basis of their 
    relatively equal frequency in adult speech directed to male and female children.

@@

Researchers, clinicians, and parents have long been fascinated with the surprising speed and variability in the growth of young children's vocabulary.
Children's early vocabulary growth is assumed to reflect not only their exposure to child-directed speech, but also the varying difficulty of different types of words, and individual differences in the aptitude of the child -- including potential language deficits.
Children show both consistency in some skills across development, as well as significant influence from external factors.
For example, @bornstein2016stability found stability in core language skills across 10 years of children's development, despite changes in maternal income and education over the study period.
Yet maternal education, often used as a proxy for socioeconomic status (SES), has also been found to be associated with children's language processing and vocabulary by 18 months [@fernald2013ses], and to be predictive of later educational outcomes [@marchman2008speed; see @schwab2016 for a review].
Demographic factors are also predictive of language skill: first-born children tend to outpace their siblings, and female children tend to have better language skills than their age-matched male counterparts [@fenson1994] -- a sex-based verbal advantage that continues through high school [see @petersen2018gender for a review].

However, it is also difficult to get a complete measure of young children's language skill: long recordings are prohibitively difficult to collect and transcribe [see @roy2015predicting, the exception that proves the rule], and yet any short recording (e.g., a 1-hour play session) will elicit only a small proportion of the words and constructions that children know. 
Thus, researchers of early word learning have constructed tests with hundreds of words, with an intentional bias to select words that are more likely to be known by children.

@@

While it is quite reasonable to expect that some CDI words are "easier" (i.e., more likely to be known by children) than others, and even to use these varying difficulties to predict variation in children's language ability, the use of psychometric models highlights the possibility that some CDI items may function differently (i.e., be more/less widely known) for different groups of children.
The idea that some items on a test may show bias, favoring one group over another, is known as Differential Item Function (DIF; @holland1993differential]. 
On any given test, it is clearly undesirable to have more items favoring one group (say, children from rural households) over another (urban children), as the test will overestimate the ability of test-takers in the former (rural) group -- and not because of any underlying mean difference in ability between the groups, but simply because the test is unfair [@camilli2013ongoing].
A variety of statistical methods for detecting DIF have been proposed, and investigations have in several instances identified DIF for many items on tests favoring one group over another (e.g., rural vs. urban).
Our goal here is to test the items on the American CDI:WS for DIF along three main axes: sex (male vs. female), maternal education (no more than secondary vs. at least some college), and race (white vs. non-white).

@@

```{r, echo=F}
sex_tab <- table(d_demo$sex)
ses_tab <- table(d_demo$ses_group)
eth_tab <- table(d_demo$ethnicity)
eth2_tab <- table(d_demo$eth_group)
momed_tab <- table(d_demo$mom_ed)
```

We analyze parent-reported Wordbank data from `r nrow(d_demo)` American English CDI: Words & Sentences administrations for children `r min(d_demo$age)` to `r max(d_demo$age)` months of age [@frank2017; @frank2021].
Full demographic data are not reported in some datasets contributed to Wordbank: sex was available for `r nrow(d_demo)-sum(is.na(d_demo$sex))` children, race/ethnicity was available for `r nrow(d_demo)-sum(is.na(d_demo$ethnicity))` children, and maternal education (a proxy for socioeconomic status; SES) was available for `r nrow(d_demo)-sum(is.na(d_demo$ses_group))` children.

The analysis of sex-based differences included CDI administrations from `r sex_tab["Female"]` female and `r sex_tab["Male"]` male children.
The analysis of race-based differences included data from `r eth_tab["White"]` white, `r eth_tab["Asian"]` Asian, `r eth_tab["Black"]` Black, `r eth_tab["Hispanic"]` Hispanic, and `r eth_tab["Other"]` "Other" children. 
Due to sparse data for many categories, we binarized participants' race/ethnicity as White (`r eth2_tab["White"]`) or Non-white (`r eth2_tab["Nonwhite"]`), recognizing that there may be important variation between groups that this will fail to capture. 
Data for the maternal education analysis included CDI data from children whose mother's had the following levels of education: `r momed_tab["Primary"]` with no more than primary school education, `r momed_tab["Some Secondary"]` with some secondary school, `r momed_tab["Secondary"]` with no more than secondary school, `r momed_tab["Some College"]` with some college, `r momed_tab["College"]` with no more than a college degree, `r momed_tab["Some Graduate"]` with no more than some graduate school, and `r momed_tab["Graduate"]` with a graduate degree.
Again due to data sparsity, we binarized the `r ses_tab["high"]` children whose mothers had at least some college or more as high maternal education (high-ME), and those whose mothers had at most high school (`r ses_tab["low"]` children) as low maternal education (low-ME).

@@

In the multigroup Rasch model, an item's difficulty is allowed to vary depending on which group an individual child belongs to. 
For example, in the sex-based multigroup model, item $i$'s difficulty is $b_{i}^{female}$ for females, and $b_{i}^{male}$ for males.
In the quest to identify DIF, a multigroup Rasch model will be fitted for each demographic dimension of interest (sex, maternal education, and ethnicity), and we will examine the between-group difficulty difference for each item (e.g., $d_i = b_{i}^{female} - b_{i}^{male}$).
If there is no DIF for a given item, then $d_i \approx 0$ as the two groups find the item equally difficult.

@@

We will first examine the size of demographic effects on language ability in a baseline Rasch model fitted without regard to demographic group.
We then fit a multigroup Rasch model for each examine evidence for DIF, assuming that 
Finally, we prune the CDI of items showing varying degrees of DIF

@@

To contextualize the later DIF results, we first fitted a baseline Rasch model to the entire dataset, without demographic information.
Figure 1 shows children's language ability vs. age by demographic group from the baseline Rasch model, which assumes no DIF (i.e., equal item parameters for all groups).
A linear regression for each demographic group, with age (centered) and its interaction, showed significant effects in the expected directions, as can also be seen in the figure. 
Female children had higher language ability than male children ($\beta=0.56$, $p<.001$), with no significant interaction with age ($\beta=0.02$, $p=.10$).
High-SES children had higher language ability than low-SES children ($\beta=0.23$, $p=.02$), an advantage that grew with age ($\beta=0.11$, $p<.001$).
White children had higher language ability than non-white children ($\beta=0.50$, $p<.001$), an advantage that grew with age ($\beta=0.08$, $p<.001$).
We will re-examine these demographic regressions after trimming items showing extreme DIF.

@@

The first step we take to identify CDI items with DIF is to create GLIMMERs (Graphs of Logits Imputed Multiply with Means Equal; @stenhaug2021treading), which visualize between-item variation in group performance differences.
These parameters are drawn from a fitted multigroup Rasch model for each demographic variable (sex, SES, and race), with the assumption that the mean language ability in each group is the same (e.g. for sex, $\mu_{male}=\mu_{female}=0$), thus pushing all between-group variation into the item difficulty parameters.
For the case of sex, where we believe $\mu_{female}>\mu_{male}$, this means we may expect to find many items with difficulty $b_i^{female} < b_i^{male}$, but we can still examine the distribution of differences in item difficulty ($d_j = b_i^{male} - b_i^{female}$) for outliers.
To give analysts a sense of the uncertainty about the existence of DIF, GLIMMER plots show distributions of parameter differences rather than point estimates. 
These distributions are generated by drawing 10,000 imputations from the item parameter covariance matrix.

Figures 2-4 show GLIMMERs for a selection of CDI items for sex, SES, and race, respectively.
The full GLIMMERs, with all 680 CDI:WS items, are available on [OSF](https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a), but were too large to include here.
Nonetheless, it is important to inspect the full plots, for if there is a cluster of items in a GLIMMER, the analyst may conclude that these items are strong candidates for DIF on that dimension.
For example, there is some clustering at the top and bottom of Figure 2: at the top, "vagina", "tights", "dress (object)", and "doll" form a cluster of items that are much easier for females, while at the bottom, "penis" stands out as much easier for males.
For the maternal education and race GLIMMERs (Figures 3 and 4)--and in the rest of the full sex GLIMMER, there are not clusters, but rather a continuum of smoothly varying differences with overlap. 
This makes identifying items with DIF quite difficult, as different methods are likely to yield inconsistent results [@stenhaug2021treading].
Nonetheless, we describe the distributions of item-level group difficulty differences, and identify extreme values, flagging items with difficulty differences that were outside 2 standard deviations of the mean of that distribution.

@@

# varying thresholds
prune_dat <- tibble()
for(thresh in seq(.25, 2.5, .25)) {
  tmp_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=thresh) # min: -.53 max: 1.06
  tmp_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=thresh) # min: -1.07 max: 1.13
  tmp_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=thresh) # min: -.64 max: 1.56
  prune_defs <- intersect(intersect(subset(tmp_ses, extreme)$definition, 
                       subset(tmp_eth, extreme)$definition),
                       subset(tmp_sex, extreme)$definition)

  # prune the columns
  keep_cols <- which(!is.element(names(d_mat), prune_defs))
  #names(d_mat)[which(is.element(names(d_mat), prune_defs))]
  d_mat_pruned <- d_mat[,keep_cols]

  m0 <- mirt(d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(-data_id), 1, "Rasch", 
           technical=list(NCYCLES=1000))

  #coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  #  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
  fscores_1pl <- tibble(data_id = subset(d_mat_pruned, rowSums(d_mat_pruned)!=0)$data_id, 
                               ability_pr = fscores(m0, method = "MAP")[,1])

  d_demo_tmp <- d_demo %>% left_join(fscores_1pl) %>%
    mutate(age_sc = scale(age, center=T, scale=F))
  # join with d_demo, do the demo regressions again..

  eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo_tmp))$coefficients # R^2=.48 
  ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo_tmp))$coefficients # R^2=.56
  sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo_tmp))$coefficients # R^2=.59

  prune_dat = bind_rows(prune_dat, 
                        tibble(SDthreshold=thresh, 
                               extreme_sex=nrow(subset(tmp_sex, extreme)),
                               extreme_ses=nrow(subset(tmp_ses, extreme)),
                               extreme_eth=nrow(subset(tmp_eth, extreme)),
                               extreme_total = length(prune_defs),
                               sex_beta = sex_reg_pr[3,1],
                               ses_beta = ses_reg_pr[3,1],
                               eth_beta = eth_reg_pr[3,1]))
}

# 2SD threshold
mm_sex2 <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06
mm_ses2 <- get_extreme_item_difficulty_differences(mm_ses) # min: -1.07 max: 1.13
mm_eth2 <- get_extreme_item_difficulty_differences(mm_eth) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex2, extreme),
                     subset(mm_ses2, extreme),
                     subset(mm_eth2, extreme)) 
# length(unique(all_extreme$definition)) 
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

@@

### Sex
For sex, the median difficulty difference (male-female) was `r round(median(mm_sex$d_diff),2)` (M=`r round(mean(mm_sex$d_diff),2)`, sd=`r round(sd(mm_sex$d_diff),2)`), with `r length(which(mm_sex$d_diff>0))`/680 items being easier for females than males.
The fact that the bulk of this distribution favors females seems to confirm that the female language advantage is pervasive across the CDI:WS, and is thus likely to be a real advantage.

### Maternal Education
For maternal education (ME), the median difficulty difference (low-high) was `r round(median(mm_ses$d_diff),2)` (M=`r round(mean(mm_ses$d_diff),2)`, sd=`r round(sd(mm_ses$d_diff),2)`), with `r length(which(mm_ses$d_diff>0))`/680 items being easier for high-ME than low-ME children. 
With the mean and median difficulty differences close to 0, and roughly half of the words favoring each ME group, it is tempting to conclude that the CDI:WS items are somewhat balanced with respect to ME.

### Race
For race, the median difficulty difference (nonwhite-white) was `r round(median(mm_eth$d_diff),2)` (M=`r round(mean(mm_eth$d_diff),2)`, sd=`r round(sd(mm_eth$d_diff),2)`), with `r length(which(mm_eth$d_diff>0))`/680 items being easier for white than non-white children, revealing a fairly pervasive advantage for white children on CDI items.

This identified 25 items with extreme sex-based difficulty differences, only 7 of which were easier for females, with the other 18 items favoring males. <!-- thus, removing these extrema would actually lower male children's vocabulary scores..] -->
For SES, 39 extrema were identified, only 13 of which were easier for low-SES children.
For race, 31 extrema were identified, only 11 of which were easier for non-white children.
Thus, while if anything the bulk of the extrema in the sex-based model advantaged the lower language ability group (males), the extrema in the other two models mostly favored the advantaged groups: 26/39 extreme items in the SES model favored high-SES children, and 20/31 items in the race model favored white children.
17 CDI items were identified as extrema in more than one model: "tractor" and "vroom" were extreme in all three models, "choo choo" was extreme in both SES and sex models, "give me five!" was extreme in both race and sex models, and 13 other items were extreme in both SES and race models ("grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy", "pet's name", "up", and "so").
In our final analysis, we explore the effect of pruning the 15 items that were extreme in both SES and race models.

@@

## The difficulty of DIF
